\section{Toplogical Groups}

\subsection{Basic properties}
Group and Topological space is two different objects, group is algebraic structure, topological space is geometric structure, but it is not strange to combine them together.

\begin{definition}
A \textbf{topological group} is a object \((G,\mathcal{T},\cdot)\) such that
\begin{itemize}
    \item \((G,\cdot)\) is a group
    \item \((G,\mathcal{T})\) is a topological space
    \item The group structure is compatible with topological structure, i.e. multiplication 
    \[m: G \times G \to G, \quad (x,y) \mapsto x \cdot y\]
    and inverse \[i: G \to G, \quad x \mapsto x^{-1}\]
    are continuous maps.
\end{itemize} 
\end{definition}

With the basic definition, we can talk about the properties of topological groups, here is a useful lemma, it shows that the topological group is \textbf{homogeneous}, i.e. the local topology is same at any point.
\begin{lemma}
    Let \(G\) be a topological group, then for any \(g \in G\), we define \textbf{left translation} and \textbf{right translation} by
    \begin{align*}
    L_g: G \to G, \quad x \mapsto gx\\
    R_g: G \to G, \quad x \mapsto xg
    \end{align*}
    Then both \(L_g\) and \(R_g\) are homeomorphisms on \(G\).
\end{lemma}
\begin{proof}
    It is easy to verify that \(L_{gh} = L_g \circ L_h\), and \(R_{gh} = R_h \circ R_g\), then we have
    \[L_g^{-1} = L_{g^{-1}}, \quad R_{g}^{-1} = R_{g^{-1}}\]
    and we can check the continuity of \(L_g\) and \(R_g\) by composition.
\end{proof}

It is an important in the following proofs of properties, we first talk about the basic properties of topological groups.

\begin{colorproposition} \label{openiscloesd}
    Let \(G\) be a topological group, then\\

    (1) each open subgroup \(H < G\) is also closed.\\

    (2) each closed subgroup \(H < G\) with finite index is also open.

\end{colorproposition}

\begin{proof}
    The subgroup \(H\) gives a partition of \(G\) by left cosets:
    \[G = \bigsqcup_{g \in \mathcal{S}} gH\]
    with \(\mathcal{S} \subset G\) the set of representatives in \(G/H\), and we fix the representatives for \(H\) be \(e\), then
    \[G - H = \bigsqcup_{g \in S-\{e\}}gH\]
    If \(H\) is open, then by translation each \(gH\) is also open, thus the union of open sets \(G-H\) is also open, hence \(H\) is closed, which finishes that proof of (1). For the statement (2), the finite index shows that \(\mathcal{S}\) is finite, and we suppose that the index is \(n\). If \(n=1\), then \(H = G\) is open clearly; otherwise, we can deduce that \(G-H\) is the finite union of closed sets \(gH\) with \(g \in \mathcal{S}-\{e\}\), thus it is closed, it can be done similarly by translation, hence \(H\) is open sice \(G-H\) is closed.
\end{proof}


Then it is the \textbf{separation} of topological groups.
\begin{colorproposition}
    Let \(G\) be a topological groups, then the following statements are equivalent:\\

    (1) \(G\) is Hausdorff (T2).\\

    (2) The singleton \(\{e\}\) is closed in \(G\).\\

    (3) \(G\) is T1.
\end{colorproposition}

\begin{proof}
    (3) implies (2) is clear, since T1 space makes all singletons closed; (1) implies (3) is just the implication from strong to weak separation axiom.\\
    
    (2) implies (1): \(G\) is Hausdorff if and only if the diagonal set
    \[\Delta = \{(g,g) \in G \times G \mid g\in G\}\]
    is closed under the product topology, and topolgical group induces a continous map \(f(x,y) = xy^{-1}\) for any \((x,y) \in G \times G\), then \(\Delta = f^{-1}(\{e\})\) is closed by continuity.
\end{proof}

\begin{corollary}
    Let \(G\) be a topological group and \(H < G\) be a subgroup, then the quotient space \(G/H\) is \textbf{Hausdorff} if and only if \(H\) is \textbf{closed} in \(G\).
\end{corollary}

Another topological properties is about \textbf{connectness}:
\begin{colorproposition}
    Let \(G\) be a topological group, then \\

    (1) If \(G\) is connected, then any open set containing \(e\) generates \(G\).\\

    (2) The connected component \(G_0\) of identity \(e \in G\) is a \textbf{closed normal subgroup} of \(G\), and each connected component of \(G\) is homeomorphic to \(G_0\) by translation.
\end{colorproposition}

\begin{proof}
    (1) We denote \(\langle A \rangle\) be the subgroup generated by \(A \subset G\), and we take \(A\) be an open set containing \(e\), we will prove that \(\langle A \rangle\) is open and closed, by the connectness of \(G\), which implies that \(\langle A \rangle = G\). It is clear that
    \[\bigcup_{x \in \langle A \rangle} xA \subset \langle A \rangle\]
    conversely, for any \(a \in \langle A \rangle\), we have \(a = ae\) with \(e \in A\), so \(a\) is in the union such that we can replace the inculsion by equality. Notice that \(xA = L_x(A)\) is open since \(A\) is open, so \(\langle A \rangle\) is also open as the union of open sets; By the propoistion \ref{openiscloesd}, \(\langle A \rangle\) is also closed since it is open subgroup, hence we finish the proof of (1).\\

    (2) Any connected component is closed in general topology, which can be proved by contradition to \(G_0 \subsetneq \overline{G_0}\). To prove that \(G_0\) is a subgroup, we take a continous map \(f(x,y) = xy^{-1}\) on \(G \times G\), then \(f(G_0 \times G_0)\) is connected since the finite product of connected spaces is connected, and we notice that \(a = f(a,e)\) for any \(a \in G_0\), so \(G_0 \subset f(G_0 \times G_0)\) and then we can take equality by the maximality of connected component, and it shows that \(G_0\) is exactly a subgroup. 

    To prove the normality, we take the conjugation map for any \(g \in G\):
    \[\sigma_g: G \to G, \quad x \mapsto gxg^{-1}\]
    it is homeomorphism by composition \(\sigma_g = L_g \circ R_g^{-1}\). Here \(\sigma_g(e) = e\) and the image is also connected, so \(\sigma_g(G_0) \subset G_0\) by the maximality of connected component, so it must be normal.
\end{proof}

Notice that if \(H < G\) is a subgroup, then it induces a natural equivalence relation on \(G\) by left cosets:
\[x \sim y \iff x=yh, \text{ for some } h \in H\]
similarly we can define right cosets here, and then it gives a quotient space \(G/\sim\) and we denote it by \(G/H\), and it is natrually endowed with the quoitent topology with a (continous) quoitent map:
\[\pi: G \to G/H, \quad g \mapsto gH\]
\textbf{In particular}, if \(H\) is a normal subgroup, then \(G/H\) is exactly a topological group by UPQ:
\[
\begin{tikzcd}
G \times G \arrow[r, "f"] \arrow[d, "\pi \times \pi"'] 
& G \arrow[d, "\pi"] \\
G/N \times G/N \arrow[r, "f_{G/N}"] 
& G/N
\end{tikzcd}
\]
with \(f(g,h) = gh^{-1}\), so it gives the continuity of group opreation. However, it is not so easy to talk about the the topology of quoitent space \(G/H\) directly, it is better to see the quotient space as a \textbf{homogeneous space} under the action of \(G\), hence we need to talk about acion of the topological group:

\begin{definition}
    Let \(G\) be a topological group and \(X\) be a topological space, a continous action of \(G\) on \(X\) is a well-defined action in the sense of group together with a continous operation, that means there exists a continous map
    \[\phi: G \times X \to X, \quad (g,x) \mapsto g \cdot x\]
    such that for any \(g,h \in G\) and \(x \in X\), we have\\

    (1) \(e \cdot x = x\)\\

    (2) \(g \cdot (h \cdot x) = (gh) \cdot x\)
\end{definition}

In the context of toplogical groups or Lie groups, we alwasys omit "continous" when it comes to actions, it is a convention sometimes in the liter

ature. It is similar to the algebraic action, we can define \textbf{stabilizer} of a point \(x \in G\)
\[I_x = \{g \in G \mid g \cdot x = x \}\]
It is sometimes called \textbf{isotropy group} at \(x\), and it can be written as the preimage
\[I_x = \phi_x^{-1}(\{x\}), \quad \phi_x(g) = g \cdot x\]
with \(\phi_x\) a continous map by fixing variable, hence the isotropy group is always a \textbf{closed subgroup} of \(G\) if \(X\) is Hausdorff (T1 is enough, but manifold makes sense). Similarly, we can define \textbf{orbit} of \(x\) by
\[Gx = O_x = \{g \cdot x \mid g \in G\} \]
which is the image of \(\phi_x\) on the contrary, hence it gives a natural bijection

\begin{lemma}[stabilizer-orbit] $ \\$
    Let \(G\) be a compact group and \(X\) be a Hausdorff space, If \(G\) acts continously on \(X\), then for any \(x \in G\), there exists a homemomorphism \(G/I_x \cong Gx\).
\end{lemma}
\begin{proof}
    The proof is just based on the universal property of quotient space, and the stablizier-orbit theorem in group theory, we just need to check the natural bijection induced by \(\phi_x\) is indeed a homemomorphism.
\end{proof}

In particular, the orbit gives a equivalence reliation on \(X\) by
\[x \sim y \iff y = g \cdot x, \text{ for some } g \in G\]
hence it gives a quotient space denoted by \(X/G\) and we call it \textbf{orbit space}, it is natrually endowed with the quoitent topology with a (continous) quoitent map:
\[\pi: X \to X/G, x \mapsto Gx\]
the following lemma is useful and important, it shows that the topology of group action is well-behaved from the view of techinc:

\begin{colorlemma}
    The natural quotient map \(\pi: X \to G/X\) is open.
\end{colorlemma}
\begin{proof}
    Suppose that \(U\) is an open set of \(X\), then we can prove
    \[\pi^{-1}(\pi(U)) = \bigsqcup_{g \in G} gU\]
    and then by translation each \(gU\) is open, so we can finish the proof by the definition of quotient topology that \(\pi(U)\) is open if its preimage is open.
\end{proof}

Then we can conclude

\begin{colorproposition}
    Let \(G\) be a topological group and \(H\) be a subgroup, then\\

    (1) \(H\) is closed if and only if \(G/H\) is Hausdorff.\\

    (2) If \(G\) is connected, then \(G/H\) is connected.\\

    (3) If \(G/H\) and \(H\) are both connected, then \(G\) is connected.
\end{colorproposition}

\begin{proof}
    
\end{proof}


\subsection*{Neighborhood system}

As we have mentioned before, topological group is a homogeneous space, so the local topology at one point can be translated to generate the global topology, for some convenience, we always study the local topology at identity of the group.

Formally, for any group \(G\) we define the \textbf{neighborhood system} (at \(x\)) to be the collection of subsets containing \(x\), denoted by \(\mathcal{V}_x\):
\[  A \in \mathcal{V}_x \iff \quad x \in A \land A \subset G\]
Hence we can induce a topology on \(G\) by neighborhood system (i.e. determine the open voisinage of identity).
\begin{proposition}
    Let \(G\) be a group with \(\mathcal{V}_e\) as the neighborhood system at identity \(e\), if \(\mathcal{V}_e\) satisfies the following conditions:
    \begin{enumerate}
    \item For any $U,V\in\mathcal{V}_e$, there exists $W\in\mathcal{V}_e$ such that 
    \[
    W \subset U\cap V.
    \]

    \item If $a\in U\in\mathcal{V}_e$, then there exists $V\in\mathcal{V}_e$ such that 
    \[
    Va \subset U.
    \]

    \item For each $U\in\mathcal{V}_e$, there exists $V\in\mathcal{V}_e$ such that 
    \[
    V^{-1}V \subset U.
    \]

    \item For each $U\in\mathcal{V}_e$ and each $x\in G$, there exists $V\in\mathcal{V}_e$ such that 
    \[
    x^{-1} V x \subset U.
    \]
\end{enumerate}
then there exist a unique topology \(\mathcal{T}\) such that \(\mathcal{V}_{e}\) is the set of all open voisinage of \(e\), and \((G,\mathcal{T})\) is a topological group.

\begin{proof}
    The topology \(\mathcal{T}\) is uniquely defined by translation:
    \[\mathcal{B} = \{gU \mid g \in G, U \in \mathcal{V}_e\}\]
    we just need to prove that \(\mathcal{B}\) is the topology basis. Notice the condition (1) and (2) ensure that \(\mathcal{B}\) generates a topology, the condition (3) is the requirement of continuity of multiplication and inverse; finally, the condition (4) is obligator for non-abelian group to ensure the continuity.
\end{proof}
\end{proposition} 
In partiuclar, we can naturally choose subgroup as the element in the system, the condition can be simplified as following:
\begin{corollary}
    Let \(G\) be a group with \( \mathcal{U}\) as a collection of subgroups of \(G\), if \( \mathcal{U}\) satisfies the following conditions:
    \begin{enumerate}
    \item For any $U,V\in\mathcal{U}$, there exists $W\in\mathcal{U}$ such that 
    \[
    W \subset U\cap V.
    \]

    \item For each $U\in\mathcal{U}$ and each $x\in G$, there exists $V\in\mathcal{U}$ such that 
    \[
    x^{-1} V x \subset U.
    \]
\end{enumerate}
then there exist a unique topology \(\mathcal{T}\) such that \(\mathcal{U}\) is a basis of open voisinage of \(e\), and \((G,\mathcal{T})\) is a topological group.
\end{corollary} 

Here are some classic examples of topological groups constructed by neighborhood system:
\begin{example}
    We consider the usual topology on \(\qq\), then the completion of \(\qq\) is just \(\rr\), it is another method to construct real number (see in Bourkabi's book). We review that we always write the viosinage of \(0\) as the form of \((-\varepsilon, \varepsilon)\), which motivates us to consider the system of subgroup:
    \[\mathcal{U} = \{\frac{1}{n}\zz \mid n \in \nn^*\}\]
    pay attention to the choices here! We can not take \(n\zz\) to be an open set (consider it in \(\rr\), it is closed and not open). Then we can verify the topology induced by \(\mathcal{U}\) is just the subspace topology induced by \(\rr\), i.e. it induces a usual absolute value on field \(\qq\):
    \[|x|_{\infty} = \begin{cases}
        x \quad x\geq 0 \\
        -x \quad x \leq 0
    \end{cases}\]
\end{example}
It is natural to ask that can we define other topological stucture on \(\qq\) such that we can get other different completion? The answer is yes, and a classic example is the \textbf{\(p-adic\) number field.}
\begin{example}
    We fix a prime number \(p\) and then we can define a subgroup 
    \[\zz_{(p)} = \{\frac{m}{n} \mid m,n \in \zz, p \nmid n\}\]
    and we let \(\mathcal{U}= \{p^t\zz_{(p)} \mid t \in \zz\}\) to be the collection of subgroups, which generates an unique toplogy (verify!).
    Notice that we have a unique chain in the system:
    \[\dots p^2 \zz_{(p)} \subset p \zz_{(p)} \subset \zz_{(p)} \subset p^{-1} \zz_{(p)} \subset p^{-2} \zz_{(p)} \subset \dots\]
    motivates from the above example, we consider each subgroup gives a method to measure the size of an elemnt is some sense, for example we consider the following number when \(p=3\):
    \[1, \quad 4, \quad \frac{1}{2}, \quad \frac{2}{5}\]
    they are all in \(\zz_{(p)}\), and we consider the distance of them to \(0\) can be bounded by \(\alpha\) (consider (\(-\varepsilon,\varepsilon\)) gives the set of all elements with distance to \(0\) less than \(\varepsilon\)), and then we consider some other numbers:
    \[\frac{1}{3}, \quad \frac{2}{3}, \quad \frac{7}{6}\]
    they are all in \(\frac{1}{3}\zz_{(p)}\) but not in \(\zz_{(3)}\), so we can think that their distance to \(0\) is bouned by \(\beta\), and it is natural to give a realtion \(\alpha < \beta\), which motivates us to give a new method to consider the metric on \(\qq\) by conisder the local factorization of \(p\).\\

    Hence we can define something to denote \(\alpha\) and \(\beta\) just now:
    \[v_p: \zz-\{0\} \to \nn, \quad m \mapsto \max\{k : p^k | m\}  \]
    for example \(v_3(6) = 2\), \(v_3(-2) = 0\). Then the method of measuring can be extended to \(\qq\) as following:
    \[v_p: \zz-\{0\} \to \zz, \quad \frac{m}{n} \mapsto v_p(m)-v_p(n)\]
    And we make a convention that \(v_p(0) = +\infty\), it is called \textbf{p-adic valustion}. for example \(v_3(\frac{1}{3}) = -1\), \(v_3(\frac{9}{2}) = 2\). Then we can get a measure of number
    \[p^k \zz_{(p)} \iff \{x \in \qq \mid v_p(x) \geq p^k\}\]
    but a question occurs here is that \(v_3(1) = 0\) and \(v_3(\frac{1}{3}) = -1\), as what we discussed just now, actually we hope the measure of \(1\) should be less than the measure of \(\frac{1}{3}\) (here \(0>-1\)), hence the valuation \(v_p\) can not be treated as the absolute value directly, hence we modify it as following:
    \[|-|_p: \qq \to \rr_+, \quad x \mapsto \frac{1}{e^{v_p(x)}}\]
    the convention just now forces \(|0|_p = 0\), and again we can get a clear measure of numbers:
    \[x \in p^k\zz_{(p)} \iff |x|_p \leq e^{-k}\]
    hence \(|1|_3 < |\frac{1}{3}|_3\), which admits our original idea.
\end{example}

\begin{remark}
    it is just a short introduction of p-adic number from the view of topologyical group, and here are sommething to supplement if to go deeper:\\

    (1) The metric of is induced by the absolute value:
    \[d(x,y) = |x-y|_p\]
    we can verify that is a \textbf{ultrametric (non-archimedean metric)}, i.e. a metric satifying the strong triangle inequality:
    \[d(x.z) \leq \max{d(x,y),d(y,z)},\quad \forall x,y,z \in \qq\]
    which induces different analysis structure on \(\qq\).\\

    (2) Same situation with \(\qq\) equipped by the usual absolute value, \((\qq,|-|_p)\) is not complete as a metric space, we can complete it to get the \textbf{p-adic  number field} \(\qq_p\). (The completion process is same with what we do for the real number, and we can prove that the completion is unique up to isometry.)\\

    (3) By \textbf{Ostrowski's theorem}, any non-trival absolute value on \(\qq\) is equivalent to either the usual absolute value or some \(p\)-adic absolute value. In detail, if \(|-|\) is a non-trival absolute value on \(\qq\), then either there is some \(c>0\) such that
    \[|x| = |x|_{\infty}^c, \quad \forall x \in \qq\]
    or there is some prime \(p\) and some \(c>0\) such that
    \[|x| = |x|_{p}^c, \quad \forall x \in \qq\]

\end{remark}

\subsection{Homogeneous Spaces}

Review that a group action \(G \circlearrowleft X\) is \textbf{transtive} if for any \(x,y \in X\), there exists \(g \in G\) such that \(g \cdot x  = y\), i.e. there is only one orbit.

\begin{definition}
    A \textbf{homogeneous space} is a Hausdorff topological space \(X\) with a transtive continous action of a group \(G\), in particular, we call it \textbf{G-homogeneous space} to emphasize the acting group \(G\).
\end{definition}

Before talking about the properties of homogeneous space, we consider a basic example but important example: rotatio action on sphere \(SO(n) \circlearrowleft \sph^{n-1}\)
\[SO(n) \times \sph^{n-1}, \quad (A,v) \mapsto Av\]
the operation is defined by matrix multiplication, so it defines a continous action, then we can verify the following results:
\begin{enumerate}
    \item \textbf{The action is transitive.} When \(n = 2\), it is clear since \(SO(n)\) is just the set of rotations of plane, so we prove the case \(n \geq 3\) by induction. We assume that the action is transitive for \(n-1\),\dots
    \item \textbf{The isotropy group at \(e_1\) is isomorphic to \(SO(n-1)\).} Suppose that \(A \in SO(n)\) is matrix such that \(Ae_1 = e_1\), then by caculation we can veify that \(A\) is of the form
    \[
M=
\begin{pmatrix}
1 & 0_{1\times n} \\
0_{n\times 1} & A'
\end{pmatrix},
\qquad A' \in SO(n-1).
\]
    Then it is easy to establish a isomorphism and a hoemomorphism.
    \item \textbf{There exists a homeomorphism \(SO(n)/SO(n-1) \cong \sph^{n-1}\).} By the stablizer-orbit lemma, we can deduce a continous bijection \(SO(n)/SO(n-1) \to \sph^{n-1}\), so it is enough to verify that \(SO(n)\) is compact.
    \item \textbf{\(SO(n)\) is connected.} When \(n = 2\), a classic homeomorphism \(SO(2) \cong \sph^1\) shows that \(SO(2)\) is connected, then we can prove the case \(n \geq 3\) by induction.
\end{enumerate}

From this example, we can see here sphere \(\sph^n\) is a homogeneous space under the action of \(SO(n+1)\), we are very faimilar with the topology of sphere, but we are not fimilar the topology of \(SO(n)\), so homogeneous space gives us a method to understand the topology of group by the homogeneous space. Conversely, if we are fimilar with the structure of the topological group, we can understand a strange space by this method, anyway homogeneous space gives a deep relation
\[\{\text{topological group}\} \sim \{\text{space with same local structure}\}\]

\begin{colortheorem}
    Let \(G\) be a Lie group and \(X\) be a \(G\)-homogeneous space, if \(I_x\) the isotropy group at \(x \in X\), then there exists a hoemomorphism
    \[G/I_x \cong X\]
\end{colortheorem}












\newpage

\subsection{The Pontryagin Duality}


We consider the category of locally compact abelian groups:
\[\cat{\mathbf{LCA}}{\text{locally compact abelian groups } (G,+)}{\text{continuous group homomorphisms }}\]

without talking about the topology of the group, we can simiarly consider the pure algebraic structure:
\[\cat{\mathbf{Ab}}{\text{abelian group } (G,+)}{\text{ group homomorphisms}}\]
Clearly it is same as category of $\mathbb{Z}$-modules, hence it is an abelian category with tools of homological algebra, so we can choose a certain objects of abelian groups to construct a duality.

\[\hat{G}: = \Hom_{\zz}(G, \rr/\zz)\]
Clearly, \(\rr/\zz\) is an injective \(\zz\)-module, hence the Hom-functor will keep exactness. In particular, it will have a good behavior when group is finite.

\begin{colorproposition}
    If \(G\) is a finite abelian group, then \(\hat{G}\) is isomorphic to \(G\).
\end{colorproposition}
\begin{proof}
    
\end{proof}

However, the duality fails when \(G\) is infinite. For example, if \(G = \zz\), then \(\hat{\zz} \cong \rr/\zz\), but if we take dual again
\[\hat{\rr/\zz} = \End_{\zz}(\rr/\zz) \cong \prod_{p}\zz_{p}\]
clearly it is larger than \(\zz\), hence it actually fails to be a duality. What we have done is to give a exact restriction on the categroy of abelian groups to make the duality work: It depends on the topology of the group, so we need to consider the topology of the dual group firstly. Formally, we define dual group as following:

\begin{definition}
    Let \(G \in \mathbf{LCA}\), we define its \textbf{dual group} as \(G^*\) or \(\hat{G}\) by
    \[G^* := \Hom_{\mathbf{LCA}}(G, \sph^1)\]
    i.e. the group of continous group homomorphism from \(G\) to the circle group (\(\sph^1 \cong \rr/\zz\)). Among \(G^*\), the elelment are called \textbf{characters} of \(G\), and we usually denote it by 
    \[\chi: G \to \sph^1\]
    with operation defined by pointwise multiplication:
    \[(\chi_1 + \chi_2)(g) := \chi_1(g) \cdot \chi_2(g)\]
\end{definition}

Generally a Hom-functor is not necessary to keep category itself, it will maps to a category of sets, so we should verify that LCA is a good condition to make the functor closed.\\

\begin{lemma}
    Let \(G \in \mathbf{LCA}\), then its dual group \(G^*\) endowed with the compact-open topology is a locally compact abelian group, in particular
    \[V(K,\varepsilon) = \{\chi \in G^* : \chi(K) \subset U_{\varepsilon}\}\]
    for any compact subset \(K \subset G\) and \(U_{\varepsilon} = \{e^{it} \mid t \in (-\varepsilon,\varepsilon)\}\), it forms a neighborhood basis of \(0 \in G^*\).
    
\end{lemma} 

\begin{proof}
    It is easy to verify that \(G^*\) is indeed a abelian group under defined opreation, to verify it is a topological group, we need to verify the family \(\mathcal{B}\) of all \(V(K,varepsilon)\) forms a neighborhood basis of identity of \(G^*\): (1) zero character \(g \mapsto 1\) is in any set clearly; (2) \(V(K,r) \cap V(L,t) = V(K\cup L, \min(r,t))\), so it is stable under finite intersection; (3) \(-V(K,\varepsilon) = V(K,\varepsilon)\) since \(U_{\varepsilon}^{-1} = U_{\varepsilon}\); (4) \(V(K,r)+V(L,t)\) is the subset of \(V(K\cup L,r+t)\), so it is stable under opreation. Hence it forms a neighborhood system of identity by (1)-(4).

    Finally, we need to verify that \(G^*\) is locally compact by Ascoli's theorem...
\end{proof}

\begin{remark}
    In particular, we denot evalution map by
    \[\ev: G \times G^* \to \sph^1, \quad (g, \chi)\mapsto \chi(g)\]
    the compact-open topology is \textbf{the coarsest topology} such that the evaluation map is continuous (page 43 [1]).
\end{remark}

A convention is that \(G^*\) always means the dual group with compact-open topology, so we finish the level of objects in the category, and then it is naturally to consider the functoriality, or the level of morphisms.

\begin{lemma}
    Let \(G,H \in \mathbf{LCA}\), and \(f: G \to H\) be a continuous group homomorphism, then it induces a natrual map
    \[f^*: H^* \to G^*, \quad \chi \mapsto \chi \circ f\]
    which is a continuous (under the compact-open topology) group homomorphism.
\end{lemma}

\begin{proof}
    For any two characters \(\chi_1, \chi_2 \in H^*\), we have
    \[f^*(\chi_1 + \chi_2)(g) = (\chi_1 + \chi_2)(f(g)) = \chi_1(f(g)) \cdot \chi_2(f(g)) = f^*(\chi_1)(g) \cdot f^*(\chi_2)(g)\]
    for any \(g \in G\), so \(f^*(\chi_1 + \chi_2) = f^*(\chi_1) + f^*(\chi_2)\) it is a group homomorphism. To prove the continuity, we take a basic open set \(V_G(K,\varepsilon)\), then
    \begin{align*}
    (f^*)^{-1}(V_G(K,\varepsilon)) &= \{\chi \in H^* \mid \chi \circ f \in V_G(K,\varepsilon)\} \\
    &= \{\chi \in H^* \mid \chi (f(K)) \subset U_{\varepsilon}\}\\
    &= V_H(f(K),\varepsilon)
    \end{align*}
    here \(f(K)\) is compact since \(f\) is continous, so the preimage of any voisinage of identity is open, hence \(f^*\) is continous by translation.
\end{proof}

It is easy to verify that \((f\circ g )^* = f^* \circ g^*\),  hence we have constructed a contravariant functor insider the category \(\mathbf{LCA}\). 

\begin{lemma}
    If \(f:G \to H\) is a surjective continuous group homomorphism, then \(f^*: H^* \to G^*\) is injective; If \(f:G \to H\) is both injective and open, then \(f^*: H^* \to G^*\) is surjective.
\end{lemma}
\begin{proof}
    If \(f\) is surjective, we take any \(\chi_1, \chi_2 \in H^*\) such that \(f^*(\chi_1) = f^*(\chi_2)\), we assume that \(\chi_1 \neq \chi_2\), then there exists \(h \in H\) such that \(\chi_1(h) \neq \chi_2(h)\), since \(f\) is surjective, there exists \(g \in G\) such that \(f(g) = h\), hence
    \(f^*(\chi_1)(g) \neq f^*(\chi_2)(g)\), so it is absurd.

    If \(f\) is injective, then injective module \(\sph^1\) allows us to extend a charcter \(\chi:G \to \sph^1\) to \(\tilde{\chi}: H \to \sph^1\) such that \(\tilde{\chi} \circ f = \chi\). We need to verify the continuity of \(\tilde{\chi}\): notice that \(f\) is open, so \(f(G)\) is the open subgroup of \(H\), hence we can prove that \(\tilde{\chi}\) is continous on \(f(G)\), then we can conclue that \(\tilde{\chi}\) is continous on \(H\) by translation (a group homomorphism is continous on some open subgroup, then the morphism is just continous).
\end{proof}

\begin{colortheorem}[Pontryagin-Van Kampen,1934] $ \\$
    Let \(G \in \mathbf{LCA}\) and \(G^*\) be its dual group, then evluation map gives a natural ismoprhism in \(\mathbf{LCA}\) by fixing one variable:
    \[\ev: G \to (G^*)^*, \quad g \mapsto \ev_g\]
    where \(\ev_g: G^* \to \sph^1\) is defined by \(\ev_g(\chi) = \chi(g)\). 
\end{colortheorem}

\begin{proof}
    
\end{proof}

Here is the slogan of the duality:
\begin{center}
    \emph{every LCA-group is the dual group of its dual group.}
\end{center}

The conclusion is similar with the finite-dimension vector space, hence it will be a strong tool to study the structure of LCA-groups and the analysis on it. In the sense of category theory, the Pontryagin duality is a contravariant equivalence with inverse itself:
\[(-)^* : \mathbf{LCA} \to ( \mathbf{LCA})^{op}\]

\subsection*{Consequences}

It is wonderful to see what we can get from the duality, we make a convention that \(G\) in this section always means a LCA-group.\\

\begin{proposition}
    In the sense of topological isomophism
    \[G_1 \oplus G_2 \cong G_1 \times G_2\]
    and 
    \[(G_1 \times G_2)^* \cong G_1^* \times G_2^*\]
\end{proposition}
\begin{proof}
    In the sense of category of abelian groups, we have
    \[(G_1 \times G_2)^* \cong G_1^* \oplus G_2^* \cong  G_1^* \times G_2^*\]
    So we just need to verify the isomophism is homemomorphism under the compact-open topology.
\end{proof}

\begin{colorproposition}
    If \(G\) is compact, then \(G^*\) is discrete; If \(G\) is discrete, then \(G^*\) is compact.
\end{colorproposition}


\begin{proof}
    
\end{proof}

\begin{colorproposition}
    If G is compact group, then\\

    - \(G\) is metrizable if and only if \(G^*\) is countable.\\

    - \(G\) is connected if and only if \(G^*\) is torsion-free.
\end{colorproposition}
    
 \newpage

\section{Classical Groups and Lie algebras}
Matrix groups are the most important examples of Lie groups and Lie algebras, this section is some preparation for the general case.

\subsection{Exponential map}

Remember that \(M_n(K) \cong K^{n^2}\) for any field (even any ring), in particualr we consider \(K = \mathbb{C}\)  or $\mathbb{R}$, then the space of endomorphisms or matirces  is just a finite-dimensional vector space over the field, so we can define the typical norm on it:
\begin{align*}
    \|A\| &= \sqrt{\sum_{i,j} |a_{ij}|^2} \quad \text{(euclidean norm)}\\
    \|A\|_{\infty} &= \max_{i,j} |a_{ij}| \quad \text{(Max norm)}\\
    \|A\|_1 &= \sum_{i,j} |a_{ij}| \quad \text{(Sum norm)}
\end{align*}


Moreover, we can consider the norm of $M_n(K)$ from another view, we view it as the linear continous operator inside $K^n$,  hence in particular the norm of $K^n$ can induce the norm of operator:
\[ \|A\|_{op} = \sup_{x \in K^n - \{0\}} \frac{\|Ax\|}{\|x\|} \]
Anyway, the space of endomrphisms on finitie-dimensional vector space or the matrix space is still finite-dimensional, as a Banach space, all the norms on $M_n(K)$ is equivalent, with the basic knowledge of topology on it, we can talk about some whether some series in analysis makes sense or not, in particular, exponential map:

\begin{definition}
    For a matrix $A \in M_n(K)$, the exponential of $A$ is defined by
    \[\exp(A) = \sum_{k \geq 0} \frac{A^k}{k!}  \]
    or sometimes we denote it by $e^A$.
\end{definition}
The definition is analogous to the exponential map in $\mathbb{C}$, and we can verify that it is well-defined for any $A$ by
\[ \|\exp(A) \|  \leq \sum_{k \geq 0} \frac{\|A\|^k}{k!} = \exp(\|A\|) < +\infty\]
so exponential map still makes sense for linear algebra! Some algebraic properties can be conclude to handle different calculation, and they are important.

\begin{proposition} \label{prop:exp} $ \\$
    Let \(K = \cc \) or \(\rr\), then\\

    (1) \(ge^Xg^{-1} = e^{gXg^{-1}}\) for any \(g \in \mathrm{GL}_n(K)\) and \(X \in M_n(K)\).\\

    (2) If \(XY = YX\), then \(e^{X+Y} = e^X e^Y\).\\

    (3) \(e^X\) is invertible with inverse \(e^{-X}\).
\end{proposition}

\begin{proof}
    (1) It is easy to check that
    \[(gXg^{-1})^k = gX^kg{-1}\]
    thus by definition we have the equality. For (2), we can expand \(e^{X+Y}\) by definition and binomial formula, and the commutativity of \(X\) and \(Y\) allows us to rerrange the terms. For (3), we see the \(X\) as the complex matrix and then by Jordan form, we have
    \[gXg^{-1} = D(I+N)\]
    under the base change of \(g\), where \(D\) is diagonal and \(N\) is nilpotent such that \(DN = ND\), thus
    \[ge^Xg^{-1} = e^De^{I+D}\]
    with \(e^D\) is diagonal and \(e^{I+D}\) an upper-triangular matrix with non-zero diagonal entries, so we can conclude that \(e^X\) is invertible with inverse \(e^{-X}\) by the property (2).
\end{proof}

There are many other proof of the propoistion, and each of them gives a new view about the matrix group and Lie group, for example, we can consider the exponential map as the solution of ODE:
\begin{colortheorem}
    For any \(X \in M_n(K)\), the one-parameter 
    \[\Phi_X:\rr \to M_n(K), \qquad t \mapsto e^{tX}\]
    is the unique \textbf{global \(C^1\) solution} of the initial value problem of ODE:
    \[\frac{d}{dt} Y(t) = X \Phi_Y(t), \qquad \Phi_X(0) = I\]
    with \(Y\) the matrix-valued function of \(t\).
\end{colortheorem}

\begin{proof}
    The uniqueness of the solution can be proved by Cauchy-Lipschitz theorem, for the existence, we should veirfy that \(\Phi_X\) is exactly the solution: For any \(t \in \rr\)
    \[\frac{\Phi_X(t+h)-\Phi_X(t)}{h} = \frac{e^{(t+h)X} - e^{tX}}{h} = e^{tX}\frac{(e^{hX} - I)}{h}\]
    The convergence of the expoential series is uniform, which allows us to exchange the limit and the sum, thus we can deduce the derivative of \(\Phi_X\) is exactly \(X\Phi_X\), hence \(\Phi_X\) is the solution of the ODE (surely it is \(C^1\), \(\Phi_X\) is derivative so it is continous, that implies that \(X\Phi_X\) is continuous too).
\end{proof}

\begin{remark}
    For property (2) in propoistion \ref{prop:exp}, here we can consider the ODE
    \[Z'(t) = (X+Y)Z(t), \qquad Z(0) = I\]
    we can verify that \(t \mapsto e^{X+Y}\) and \(t \mapsto e^{X}e^{Y}\) are both solutions when \(X\) and \(Y\) commute, thus they are equal by the uniqueness of the solution.
\end{remark}

\begin{remark}
    Actually, the theorem contain an simple information is that
    \[\frac{d}{dt}(e^{tX}) = Xe^{tX}\]
    so we notice the amazing similarity with the case in euclidean space, which allows us to consider the one-parameter subgroup as a dynamic system, and the exponential map is the flow of the system, this is the original ideal of the exponential map in Lie theory.
\end{remark}

Another method to prove that \(e^X\) is invertible is to consider its determinant, which is natural to motivate the following result:

\begin{colortheorem}[Jacobi's formula]
    For any \(A \in M_n(K)\), we have
    \[\det(e^A) = e^{\mathrm{tr}(A)}\]
    
\end{colortheorem}
\begin{proof}
    By Schur's lemma, any \(X \in M_n(\cc)\) can be upper-triangularized by some \(g \in \mathrm{GL}_n(\cc)\) such that \(X = gTg^{-1}\) the upper-trianglue matrix \(T\) with diagonal entries \(\lambda_1,...,\lambda_n\), thus
    \[\det (e^X) = e^{\lambda_1}\cdots e^{\lambda_n} =  e^{\lambda_1+\dots+\lambda_n} = e^{\mathrm{tr}(X)}\]
\end{proof}

Another analytic aspects of the exponential map is asymptotic expansion, that means we can manuiplate the exponential map as what we do in analysis, but first of all we prove that the exponential map is similarly a smooth map by the requirement of the expansion:

\begin{proposition}
    For any \(X \in M_n(K)\), the one-parameter
    \[\Phi_X: \rr \to GL_n(K), \quad t \mapsto e^{tX}\]
    defines a real-analytic homomorphism, and we have the following expansion\\

    (1) \(\exp(tX)\exp(tY) = \exp(t(X+Y)+\frac{1}{2}[X,Y]+O(t^3))\)\\
    
    (2) \(\exp(tX)\exp(tY)\exp(-tX)\exp(-tY) = \exp(t^2[X,Y]+O(t^3))\)\\

    for any \(X,Y \in M_n(K)\), with bracket \([X,Y] = XY - YX\) the commutator of \(X\) and \(Y\).
\end{proposition}
\begin{proof}
    For any \(t_0 \in \rr\), we can expand \(\Phi_X\) at \(t_0\) by definition:
    \[e^{tX} = e^{t_0X}e^{(t-t_0)X} = e^{t_0X} \sum_{k=0}^\infty \frac{(t-t_0)^k}{k!} X^k\]
    The series converges uniformly on any compact interval, so the map is real-analytic. To prove the formula, we just need to expand the left-hand side by definition and rearrange the terms.
\end{proof}

\begin{remark}
    If \(XY = YX\), then the commutator \([X,Y] = 0\), thus the expansion in (1) is exact, which is just the property (2) in proposition \ref{prop:exp}, hence we get the same result by using the expansion.
\end{remark}

Above properties of one-parameter allows us to consider the general case in matrix group, later we will see that the matrix group is indeed a smooth manifold, i.e. a Lie group.

\begin{theorem}
    Let \(\gamma: \rr \to GL_n(\rr)\) be a one-parameter subgroup, then there exists a unique \(X \in M_n(\rr)\) such that \(\gamma(t) = e^{tX}\) for any \(t \in \rr\), so any one-parameter subgroup of \(GL_n(\rr)\) is exactly real-analytic.
\end{theorem}
\begin{proof}
    convolution...
\end{proof}

\subsection{Correspondence between lie group and lie algebra}
Formally, we define the real-general linear group as following:
\[GL_n(\rr) : = \{A \in M_n(\rr) \mid \det(A) \neq 0\}\]
it is the subset of \(M_n(\rr)\), while the topology of \(GL_n(\rr)\) is just the subspace topology induced by \(M_n(\rr) \cong \rr^{n^2}\), hence we can conlude something about the topology of the group:

\begin{itemize}
    \item \(GL_n(\rr)\) is a open subset of \(M_n(\rr)\), it is immediately by
    \[GL_n(\rr) = \mathrm{det}^{-1}(\rr-\{0\})\]
    since determinant is a continuous map, so it can be viewed as a open subset of euclidean space \(\rr^{n^2}\), so it is a smooth manifold.
    \item Following the above result, since it can be viewed as an open subset of euclidean space, so it is a "bon espace" in the sense of topology, i.e. semi-locally simply connected, locally path-connected, locally compact, and Hausdorff.
    \item Consider the opreation of the group
    \[m(A,B) = AB^{-1} = \frac{1}{\det(B)}A\cdot adj(B)\]
    where \(adj(B)\) is the adjoint matrix of \(B\), which is a polynomial of entries of \(B\), hence this map is smooth as polynomial map, which shows that \(GL_n(\rr)\) is a topological group, and in particular, with the good smooth structure, it is a Lie group.
    \item As the topological group, the local structure is same , so we can just consider the point \(I \in GL_n(\rr)\). As a differential manifold, we can consider the tangent space at \(I\), so we can identify the tangent space
    \[T_I(\mathrm{GL}_n(\rr)) \cong \rr^{n^2} \cong M_n(\rr)\]
    Actually, we can view a strong result: For any \(X \in M_n(\rr)\), we define a curve on manifold by \[\gamma_X: \rr \to GL_n(\rr), \quad t \mapsto e^{tX}\]
    by exponential map, this is smooth natrual curve with derivative \(\gamma_X'(0) = X\), hence we prove that \(M_n(\rr)\) is exactly the tangent space of \(GL_n(\rr)\) at \(I\), and we realize the connection between the local structure of the Lie group and its tangent space by one-parameter subgroup, that is the basic ideal of Lie theory.
\end{itemize}

We conclude above discussion by the following definition, or exactly the \textbf{vocabulary}, because they are just te new notation of manifold and vector space, but they are the basic object of Lie theory, so we need to give them a name:
\begin{definition} 
    A \textbf{lie group} is a smooth manifold \(G\) with a group structure such that the group operation \(m: G \times G \to G\) defined by \(m(g,h) = gh^{-1}\) is smooth.\\

    A \textbf{real lie algebra} is a \(\rr\)-vector space \(\mathfrak{g}\) with a bilinear map \([-, -]: \mathfrak{g} \times \mathfrak{g} \to \mathfrak{g}\) such that for any \(x,y,z \in \mathfrak{g}\) and \(a,b \in \rr\), we have\\

    (1) \([x,y] = -[y,x]\) (antisymmetry)\\

    (2) \([x,[y,z]] + [y,[z,x]] + [z,[x,y]] = 0\) (Jacobi identity)
\end{definition}

with above discuss we can see that \(GL_n(\rr)\) is exactly a Lie group, while its tangent space at identity \(M_n(\rr)\) is a vector space, actually it is a Lie alegbra by defining the bracket as following:
\[[X,Y] = XY - YX\]
and we say \(\mathfrak{h}\) is a \textbf{sub(-lie) algebra} of \(\mathfrak{g}\) if \(\mathfrak{h}\) is a subspace of \(\mathfrak{g}\) and is closed under the bracket, i.e.
 \[\forall x,y \in \mathfrak{h}, [x,y] \in \mathfrak{h}\]
To study other matrix groups, we can just consider the submanifold of \(GL_n(\rr)\), i.e. the subgroup of \(GL_n(\rr)\) with the subspace topology, hence we can get the following result:

\begin{colortheorem} \label{linearliegroup}
    Let \(G\) be a closed subgroup of \(GL_n(\rr)\), then \(G\) is a submanifold of \(GL_n(\rr)\), and the tangent space of \(G\) at identity is a Lie algebra with
    \[T_I(G) = \{M \in M_n(\rr) \mid e^{tM} \in G, \forall t \in \rr\}\]
    it is a \textbf{subalgebra} of \(M_n(\rr)\) with the bracket defined by commutator.
\end{colortheorem}

\begin{proof}
    
\end{proof}

which motivates the following definition:

\begin{colordefinition}
    A \textbf{linear lie group} is a closed subgroup of \(GL_n(\rr)\) for some \(n\), and its lie algebra is defined as its tangent space
    \[\mathrm{Lie}(G) = \{M \in M_n(\rr) \mid e^{tM} \in G, \forall t \in \rr\}\]
    and we usually denote it by frak character \(\mathfrak{g}\).
\end{colordefinition}

We invite the Lie algebra of \(GL_n(\rr)\) by consider its tangent space at identity, conversely, exponential map allows us to know the local structure of the group by its Lie algebra, which can be done beyond the general manifold
\begin{theorem}
    The exponential map \(\exp: M_n(\rr) \to GL_n(\rr)\) is a local diffeomorphism at \(0\), in particular, it is a diffeomorphism from some neighborhood of \(0\) to the image.
\end{theorem}

\begin{proof}
    
\end{proof}

Let us consider some examples related to the general linear group:

\begin{example}
    We view \(\cc\) as a real vector space, then there exists a natural embedding
    \[\iota:M_n(\cc) \to M_{2n}(\rr), \quad A+iB \mapsto \begin{pmatrix} A & -B \\ B & A \end{pmatrix}\]
    it is a injective continous linear map, and we can verify that 
    \[(A+iB)(x+iy) = 0 \iff \begin{pmatrix}
        A & -B \\
        B & A 
    \end{pmatrix} \begin{pmatrix} x \\ y \end{pmatrix} = \begin{pmatrix} 0 \\ 0 \end{pmatrix}\]
    which shows that restriction of \(\iota\) on \(GL_n(\cc)\) takes image in \(GL_{2n}(\rr)\), and image \(\iota(GL_n(\cc))\) is indeed a closed subgroup of \(GL_{2n}(\rr)\), hence it is a linear lie group, and its lie algebra is exactly the image of \(\iota\) on \(M_n(\cc)\), which motivates the following diagram    
    \[
\begin{tikzcd}
M_n(\mathbb{C}) \arrow[r, "\iota"] \arrow[d, "\exp"'] 
& M_{2n}(\mathbb{R}) \arrow[d, "\exp"] \\
\mathrm{GL}_n(\mathbb{C}) \arrow[r, "\iota"] 
& \mathrm{GL}_{2n}(\mathbb{R})
\end{tikzcd}
\]
similarly with general linear group, the closed subgroup of \(\GL_n(\cc)\) is also a submanifold of \(\GL_n(\cc)\), and the tangent space at identity is the subalgebra of \(M_n(\cc)\) with the bracket defined by commutator, hence we can get a large class of linear lie groups and lie algebras by this method, it is a big case of linear lie groups.
\end{example}

\begin{example} (Affine group)
    We consider a closed subgroup of \(\GL_{n+1}(\rr)\) defined by
\[G = \left\{\begin{pmatrix}A & b \\ 0 & 1\end{pmatrix} \mid A \in GL_n(\rr), b \in \rr^n\right\}\]
and we consider the group action \(G \circlearrowleft \rr^n\) defined by
\[\begin{pmatrix}
    A & b\\
    0 & 1
\end{pmatrix}\cdot x = \begin{pmatrix}
    A & b\\
    0 & 1
\end{pmatrix}\begin{pmatrix}
    x\\
    1
\end{pmatrix} = Ax +b\]
This action is transitive so \(\rr^n\) is a homogeneous space under \(G\), and we can establish a homeomorphism
\[G/\GL_n(\rr) \cong \rr^n\]
It is a classic example about Erlangen progamme, here \(G\) is a important geometric group, we usually call it the \textbf{affine group} \(\mathrm{Aff}_n(\rr)\).
\end{example}


\subsection{Adjoint map}
Review a bsic fact about the finite-dimensional vector space: any norm on a finite-dimensional Banach space is equivalent, hence it is natural to consider the abstract exponential map on linear opreators

\begin{definition}
    Let \(V\) be a finite-dimensional vector space over \(\rr\) or \(\cc\), we define the exponential map \(\Exp: \End(V) \to \GL(V)\) by
    \[\Exp(T)  = \sum_{k \geq 0} \frac{T^k}{k!}\]
    where \(T^k = T\circ \cdots \circ T\) is the \(k\)-th composition of \(T\).
\end{definition}

The definition is well-defined when we consider \(\End(V)\) as a Banach space, and we can verify that the properties in proposition \ref{prop:exp} still hold for the abstract exponential map. In particular, if we choose a basis of \(V\) and denote it by \(\mathcal{B}\), then we can see their connection by the following commutative diagram:
% https://q.uiver.app/#q=WzAsNCxbMCwwLCJcXG1hdGhybXtFbmR9KFYpIl0sWzEsMCwiXFxtYXRocm17R0x9KFYpIl0sWzAsMSwiTV9uKEspIl0sWzEsMSwiXFxtYXRocm17R0x9X24oSykiXSxbMCwxLCJcXG1hdGhybXtFeHB9Il0sWzAsMiwiW1xcY2RvdF1fe1xcbWF0aGNhbHtCfX0iLDJdLFsyLDMsIlxcZXhwIiwyXSxbMSwzLCJbXFxjZG90XV97XFxtYXRoY2Fse0J9fSJdXQ==
\[\begin{tikzcd}
	{\mathrm{End}(V)} & {\mathrm{GL}(V)} \\
	{M_n(K)} & {\mathrm{GL}_n(K)}
	\arrow["{\mathrm{Exp}}", from=1-1, to=1-2]
	\arrow["{[\cdot]_{\mathcal{B}}}"', from=1-1, to=2-1]
	\arrow["{[\cdot]_{\mathcal{B}}}", from=1-2, to=2-2]
	\arrow["\exp"', from=2-1, to=2-2]
\end{tikzcd}\]
or simply the equality \[[\Exp(T)]_{\mathcal{B}} = \exp([T]_{\mathcal{B}})\]


\textbf{Our first aim is to construct the lie alegbra from exactly a lie group.}\\

We consider the conjugation action of a linear lie group \(G\) on itself, and then for any \(g \in G\), there exists a diffeomorphism by the smooth structure of \(G\)
\[c_g: G \to G, \quad h \mapsto ghg^{-1}\]
with \(c_g(I) = I\), hence it allows us to consider the differential of \(c_g\) at identity,
\[(Dc_g)_I: T_I(G) \to T_I(G), \quad X \mapsto gXg^{-1} \]
the reason is as following: for linear lie group, the tangent space at identity is just \(\mathrm{Lie}(G)\) by theorem \ref{linearliegroup}, so for any \(X \in \mathrm{Lie}(G)\), there exists a one-parameter subgroup \(\gamma_X: \rr \to G\) defined by \(\gamma_X(t) = e^{tX}\) with \(X = \frac{d}{dt}\big|_{t=0} \gamma_X(t)\), hence we consider the parameterziation of the conjugation
\[\frac{d}{dt}\big|_{t=0} c_g(\gamma_X(t)) = \frac{d}{dt}\big|_{t=0} g\gamma_X(t)g^{-1} = gXg^{-1}\]
which shows that \((Dc_g)_I(X) = gXg^{-1}\) exactly, hence by the construction we induce a representation
\[\Ad: G \to \GL(\mathfrak{g}), \quad g \mapsto Ad_g = (Dc_g)_I\]
and we call the representation \(\Ad\) the \textbf{adjoint representation} of \(G\).

Notice that \(\mathfrak{g}\) is finite-dimensional, so \(\GL(\mathfrak{g})\) is also a linear Lie group, with its Lie algebra \(\End(\mathfrak{g})\), hence we can consider the differential of \(\Ad\) at identity

\begin{proposition}
    Let \(G\) be a linear lie group with lie algebra \(\mathfrak{g}\), then the direction derivative of the adjoint representation at identity is 
    \[(D\Ad)_I(X) = \frac{d}{dt}\big|_{t=0} \Ad_{\exp(tX)} = \ad_X\]
    with \(\ad(X): \mathfrak{g} \to \mathfrak{g}\) defined by \(\ad_X(Y) = XY-YX\) the commutator of \(X\) and \(Y\).
\end{proposition}

\begin{proof}
    caculation...
\end{proof}

Hence we get the basic idea of the adjoint 

\begin{definition}
    Let \(G\) be a linear lie group with its lie algebra \(\mathfrak{g}\), then 
    \begin{itemize}
        \item The adjoint representation of \(G\) is defined by
        \[\Ad: G \to \Aut(\mathfrak{g}), \quad g \mapsto \Ad_g\]
        with \(\Ad_g(X) = gXg^{-1}\) for any \(X \in \mathfrak{g}\).\\
        \item The adjoint representation of Lie algebra \(\mathfrak{g}\) is defined by
        \[\ad: \mathfrak{g} \to \End(\mathfrak{g}), \quad X \mapsto \ad_X\]
        with \(\ad_X(Y) = [X,Y]\) for any \(Y \in \mathfrak{g}\).
    \end{itemize}
\end{definition}

\begin{remark}
    An application bettween two lie algebras \(f: \mathfrak{g} \to \mathfrak{h}\) is a \textbf{morphism of Lie algebra} if it is a linear map and preserves the bracket \[f([X,Y]_{\mathfrak{g}}) = [f(X), f(Y)]_{\mathfrak{h}}\]
    In particular, the isomorphism in the categroy of lie algebras is called \textbf{automorphism}, and we use \(\Aut(\mathfrak{g})\) to denote the group of automorphisms of \(\mathfrak{g}\), which is subspace of \(\GL(\mathfrak{g})\), the latest one is the set of all bijective linear maps on \(\mathfrak{g}\). 
\end{remark}

with above discussion, we can describe a core properties of the adjoint representation, which is the connection between the local structure of the group and its lie algebra:

\begin{colortheorem}
    Let \(G\) be a linear lie group with its lie algebra \(\mathfrak{g}\), then for any \(X \in \mathfrak{g}\), we have \[\Ad_{\exp(X)} = \Exp(\ad_X)\]
    i.e. the following diagram commutes:
    \[
\begin{tikzcd}
\mathfrak g \arrow[r, "\ad"] \arrow[d, "\exp"'] 
& \End(\mathfrak g) \arrow[d, "\Exp"] \\
G \arrow[r, "\Ad"] 
& \GL(\mathfrak g)
\end{tikzcd}
\]
\end{colortheorem}

\begin{proof}
    
\end{proof}







\newpage

\section{Lie Algebra}

In this section all vector space will be finite-dimensional.

\subsection{Ideal}

\begin{definition}
    Let \(\mathfrak{g}\) be a lie algebra with bracket, then \(\mathfrak{h} \subset \mathfrak{g}\) is an ideal if \(\mathfrak{h}\) is a linear subalgebra and \([g,h] \in \mathfrak{h}\) for any \(g \in \mathfrak{g}\) and \(h \in \mathfrak{h}\).
\end{definition}

Ideal is like the ideal in ring, it allows us to talk about the quotient of Lie algebra, i.e. we consider the quotient of linear space firstly
\[\pi: \mathfrak{g} \to \mathfrak{g}/\mathfrak{h}, \quad X \mapsto X + \mathfrak{h}\]

and then \(\pi\) is a lie algebra morphism if and only if \(\mathfrak{g}/ \mathfrak{h}\) is a well-defined lie algebra, that allows us to write
\[[x + \mathfrak{h}, y + \mathfrak{h}] = [\pi(x),\pi(y)] = \pi([x,y]) = [x,y]+ \mathfrak{h}\]
which ensures us to define the bracket on the quotient linear space such that \(\mathfrak{g}/\mathfrak{h}\) is a lie algebra, and similarly we have the isomorhism theorem.

\begin{theorem}
    If \(f: \mathfrak{g} \to \mathfrak{h}\) is a lie algebra morphism, then \(\ker f\) is an ideal of \(\mathfrak{g}\) and \(\im f\) is a subalgebra of \(\mathfrak{h}\), and we have the isomorphism of lie algebra
\[\mathfrak{g}/\ker f \cong \im f\]
\end{theorem}

\begin{remark}
    Here are some quick notes about the ideal:
    \begin{enumerate}
        \item The intersection of ideals and sum of ideals are still ideals.
        \item The intersection of ideal \(I\) and subalgebra \(\mathfrak{g}\) will \textbf{be the ideal of subalgebra}, and it is necessary to be the ideal of the original algebra.
        \item Two important examples of ideal, the first one is the \textbf{centre of a lie algebra}
        \[Z(\mathfrak{g}) = \{X \in \mathfrak{g} \mid [X,Y] = 0, \forall Y \in \mathfrak{g}\}\]
        and the second one is the \textbf{derived algebra} of a lie algebra
        \[[\mathfrak{g}, \mathfrak{g}] = \{\sum_{i=1}^n c_i[x_i,y_i] \in \mathfrak{g} \mid x_i,y_i \in \mathfrak{g}, c_i \in k, n \in \mathbb{N}\}\]
        \item We can define the ideal generated by two subsets \(A\) and \(B\) by 
        \[[A,B] = \mathrm{span}\{[a,b] \mid a \in A, b \in B\}\]
        A notation should be clarified is that sometimes we write \(\mathfrak{a}\mathfrak{b}\) to instead of \([\mathfrak{a},\mathfrak{b}]\), the reason is that the bracket is just a non-associative product defined on a module.
    \end{enumerate}
\end{remark}

\begin{definition}
    Let \(\mathfrak{g}\) be a lie algebra, and we define
    \begin{itemize}
        \item The \textbf{central series} of \(\mathfrak{g}\) is defined by
        \[\mathcal{C}(\mathfrak{g}) = [\mathfrak{g},\mathfrak{g}], \quad \mathcal{C}^{n+1}(\mathfrak{g}) = [\mathfrak{g}, \mathcal{C}^{n}(\mathfrak{g})]\]
        for any \(n \geq 1\).\\

        \item The \textbf{derived series} of \(\mathfrak{g}\) is defined by
        \[\mathcal{D}(\mathfrak{g}) = [\mathfrak{g},\mathfrak{g}], \quad \mathcal{D}^{n+1}(\mathfrak{g}) = [\mathcal{D}^{n}(\mathfrak{g}), \mathcal{D}^{n}(\mathfrak{g})]\]
        for any \(n \geq 1\).
    \end{itemize}
\end{definition}

This is two types of decreasing chains of ideals, sometimes we swap the noatation as following:
\[\mathcal{C}^n(\mathfrak{g}) = \mathfrak{g}^n, \quad \mathcal{D}^n(\mathfrak{g}) = \mathfrak{g}^{(n)}\]
for any \(n \geq 1\), the reason is same as the notation of derived algebra. Another useful notation is for quotient algebra, we can write
\[\mathcal{C}^n(\mathfrak{g}/\mathfrak{h}) = \mathcal{C}^n(\mathfrak{g}) + \mathfrak{h} / \mathfrak{h}, \quad \mathcal{D}^n(\mathfrak{g}/\mathfrak{h}) = \mathcal{D}^n(\mathfrak{g}) + \mathfrak{h} / \mathfrak{h}\]


\begin{colordefinition}
    Let \(\mathfrak{g}\) be a lie algebra, then\\

    - \(\mathfrak{g}\) is \textbf{nilpotent} if there exists some \(n\) such that \(\mathcal{C}^n(\mathfrak{g}) = 0\).\\

    - \(\mathfrak{g}\) is \textbf{solvable} if there exists some \(n\) such that \(\mathcal{D}^n(\mathfrak{g}) = 0\).
\end{colordefinition}

Here are some results immediately by the definition:
\begin{enumerate}
    \item nilpotent implies solvable since \(\mathcal{D}^n(\mathfrak{g}) \subset \mathcal{C}^n(\mathfrak{g})\) for any \(n\).
    \item If \(\mathfrak{g}\) is nilpotent and non-zero, then \(Z(\mathfrak{g})\) is non-zero.
    \item If \([\mathfrak{g},\mathfrak{g}]\) is nilpotent, then \(\mathfrak{g}\) is solvable.\\
    (Hint: by definition we can prove \(\mathcal{D}^{n+1}(\mathfrak{g}) \subset \mathcal{C}^n([\mathfrak{g},\mathfrak{g}])\))
\end{enumerate}

The two properties will keep stable under the morphism of lie algebra, that is the following result:

\begin{proposition}
    Let \(\mathfrak{g}, \mathfrak{h}\) and \(\mathfrak{f}\) be lie algebras satisying the short exact sequence
    \[0 \to \mathfrak{f} \to \mathfrak{g} \to \mathfrak{h} \to 0\]
    then \(\mathfrak{g}\) is nilpotent (resp. solvable) if and only if \(\mathfrak{f}\) and \(\mathfrak{h}\) are nilpotent (resp. solvable), equivalently it is enough to prove the three properties:\\

    (1) The subalgebra of a nilpotent (resp. solvable) lie algebra is still nilpotent (resp. solvable).\\

    (2) The quotient of a nilpotent (resp. solvable) lie algebra is still nilpotent (resp. solvable).\\

    (3) If \(\mathfrak{g}/\mathfrak{h}\) and \(\mathfrak{h}\) are nilpotent (resp. solvable), then \(\mathfrak{g}\) is nilpotent (resp. solvable).
\end{proposition}

\begin{proof}
    
\end{proof}









\subsection*{upper-triangular matrix}



We have seen that many matirx ring is exactly a lie algebra, and each linear lie group matches a lie algebra, which motivates to consider the representation of an abstract lie algebra.

\begin{colordefinition}
    
\end{colordefinition}

\begin{lemma} \label{lemmaengel}
For any representation \(\rho: \mathfrak{g} \to \End(V)\) with \(V\) non-zero, if \(\rho(x)\) is nilpotent for any \(x\), then there exists some non-zero vector \(v \in V\) such that \(\rho(x)(v) = 0\) for any \(x\).     
\end{lemma}


\begin{colortheorem}[Engel's theorem]
    Let \(\mathfrak{g}\) be a lie algebra, then \(\mathfrak{g}\) is nilpotent if and only if \(\mathrm{ad}_x\) is nilpotent for any \(x \in \mathfrak{g}\).
\end{colortheorem}

\begin{proof}
    It is clearly all adjoint map is nilpotent in \(\End(\mathfrak{g})\) if \(\mathfrak{g}\) is nilpotent, so it is enough to prove the converse. If \(\mathfrak{g}= 0\) or \(\dim \mathfrak{g} = 1\), then \(\mathfrak{g}\) is nilpotent, so we can assume for any \(\dim \mathfrak{g} = n \geq 2\), if \(\mathrm{ad}_x\) is nilpotent for any \(x\), then \(\mathfrak{g}\) is nilpotent. 

    By lemma \ref{lemmaengel} there exists \(y \in \mathfrak{g}\) such that \(\mathrm{ad}_x(y) = [x,y] =0\) for any \(x \in \mathfrak{g}\), so it means \(Z(\mathfrak{g}) \neq 0\). Then we can consider the quotient algebra \(\mathfrak{g}/Z(\mathfrak{g})\) with dimension less than \(n\), and any adjoint map of \(\mathfrak{g}/Z(\mathfrak{g})\) is nilpotent if any adjoint map of \(\mathfrak{g}\) is nilpotent, hence by induction \(\mathfrak{g}/Z(\mathfrak{g})\) is nilpotent, so \(\mathfrak{g}\) is nilpotent since \(Z(\mathfrak{g})\) is always nilpotent.
\end{proof}

\begin{colortheorem}[Lie's theorem]
    Let \(\mathfrak{g}\) be a \textbf{solvable} lie algebra over an \textbf{algebraically closed field with characteristic zero}, then for any finite-dimensional representation \(\rho: \mathfrak{g} \to \End(V)\), there exists \(v \in V-{0}\) such that 
    \[\rho(x)(v)  = \lambda(x) \cdot v\]
    for some scalar function \(\lambda: \mathfrak{g} \to K\) and any \(x \in \mathfrak{g}\).
\end{colortheorem}

\begin{proof}
    
\end{proof}

\subsection{Killing form}

\begin{colordefinition}
    A \textbf{non-abelian} lie algebra \(\mathfrak{g}\) is simple if \(\mathfrak{g} \neq 0\) and the only ideals are \(0\) and \(\mathfrak{g}\).
\end{colordefinition}

\begin{remark}
    The condition of non-abelian is necessary, otherwise the definition of ideal will be reduced to the linear subspace, and the unique case of "simple" algebras is one-dimensional lie algebra.
\end{remark}

\begin{proposition}
        The lie algebra \(\mathfrak{g}\) is simple if and only if the adjoint representation \(\ad: \mathfrak{g} \to \End(\mathfrak{g})\) is irreducible.
\end{proposition}

\begin{proof}
    Suppose that the adjoint representation is not irreducible, then there exists a proper non-zero subspace \(V \subset \mathfrak{g}\) such that \(\ad_X(V) \subset V\) for any \(X \in \mathfrak{g}\), i.e.
    \[\forall X \in \mathfrak{g}, \forall Y \in V, \quad [X,Y] \in V\]
    it shows that \(V\) is an proper non-zero ideal of \(\mathfrak{g}\), hence \(\mathfrak{g}\) is not simple. Conversely, notice that if \(\mathfrak{h}\) is an ideal of \(\mathfrak{g}\), then 
    \[\forall X \in \mathfrak{g}, \forall Y \in \mathfrak{h}, \quad [X,Y] \in \mathfrak{h}\]
    i.e. which shows that \(\mathfrak{h}\) is an invariant subspace of the adjoint representation, hence if \(\mathfrak{g}\) is simple, then the adjoint representation is irreducible.
\end{proof}

\begin{colordefinition}
    The radical of a lie algebra \(\mathfrak{g}\) is defined as the maximal sovlable ideal of \(\mathfrak{g}\), and we denote it by \(\mathrm{rad}(\mathfrak{g})\).
\end{colordefinition}

\begin{remark}
    The definition is well-defined since \textbf{the sum of two solvable ideals is still a solvable ideal}, alternatively, we can define the radical ideal as 
    \[\mathrm{rad}(\mathfrak{g}) = \sum \mathfrak{h}\]
    where the sum runs over all solvable ideals \(\mathfrak{h}\) of \(\mathfrak{g}\), and the finite-dimensional condition ensures the choice of ideals is finite, hence the sum is exactly an ideal.
\end{remark}

\begin{definition}
    The killing form of a lie algebra \(\mathfrak{g}\) is the trace form of its adjoint representation, i.e. it is the bilinear form defined by
    \[\kappa(x,y) = \mathrm{tr}(\mathrm{ad}_x \circ \mathrm{ad}_y)\]
\end{definition}

\begin{remark}
    Here are some properties of the killing form:
    \begin{enumerate}
        \item The killing form is symmetric since the trace is symmetric.
        \item The killing form is invariant under the adjoint representation 
        \[\kappa([x,y],z) + \kappa(y,[x,z]) = 0\]
    \end{enumerate}
\end{remark}

\begin{proposition}
    Let \(\mathfrak{g}\) be a lie algebra, then the following statements are equivalent:
    \begin{enumerate}[(a)]
        \item  The only abelian ideal is \(\{0\}\).
        \item \(\mathrm{rad}(\mathfrak{g}) = \{0\}\) 
        \item[] \textbf{If the base field is of characteristic \(0\)}
        \item The killing form of \(\mathfrak{g}\) is non-degenerate.
        \item \(\mathfrak{g}\) is a direct sum of simple lie algebras.
    \end{enumerate}
\end{proposition}

\begin{proof}
    
\end{proof}
